### 2024-03-12 paper/paper.qmd
Add model set-up section to paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

# Data {#sec-data}

# Model

## Model set-up

The goal of our modeling strategy is to forecast if a person voted for Biden, based only on knowing their gender, race, and gun ownership status.

The model that we are interested in is:

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta \times \mbox{gender}_i + \gamma \times \mbox{education}_i + \delta \times \mbox{gun}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\delta &\sim \mbox{Normal}(0, 2.5)
\end{align}

Where:

* $y_i$ is the binary outcome variable, representing who respondent $i$ voted for and equal to 1 if Biden and 0 if Trump, 
* $\pi_i$ is the probability that respondent $i$ voted for Biden, 
* ${gender}_i$ is a predictor variable, representing the gender of respondent $i$, 
* ${race}_i$ is a predictor variable, representing the race of respondent $i$, and 
* ${gun}_i$ is a predictor variable, representing the gun ownership status of respondent $i$.

We used a logistic regression model in a Bayesian framework using the package `rstanarm` [@rstanarm], which we will briefly describe here. Logistic regression is a type of generalized linear model. It is a tool for data exploration and used when we are interested in the relationship between a binary outcome variable and some predictor variables.

The foundation of logistic regression is the Bernoulli distribution and logit function. The Bernoulli distribution is a discrete probability distribution having two possible outcomes, “1” and “0”, in which “1” occurs with probability $p$ and “0” occurs with probability $1-p$. Logistic regression is still a linear model, because the predictor variables enter in a linear fashion [@rohan]. Hence, the logit function links the Bernoulli distribution to the machinery we use in linear models [@rohan].

In our model, we also have the parameters $\alpha$, $\beta$, $\gamma$, and $\delta$ in addition to the variables. The parameter $\alpha$ is the intercept and the parameters, $\beta$, $\gamma$, and $\delta$, are the slope coefficients. We specify prior probability distributions for each of the parameters in our model, but these are just the default priors that `rstanarm` [@rstanarm] uses (Normal distribution with mean and standard deviation of 0 and 2.5).

## Model justification

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

\newpage

# References

### 2024-03-12 paper/paper.qmd
Add introduction section to paper

@@
 author: 
-  - First author
-  - Another author
+  - Tam Ly
+  - Renfrew Ao-Ieong
+  - Rahma Binth Mohammad
 thanks: "Code and data are available at: LINK."
@@
 
+The United States (US) presidential election of 2020 resulted in President Joe Biden’s victory, making him the 46th President of the US.
+The two top runners were Biden for the Democrats and former President Donald Trump for the Republicans.
+The presidential race between the two was of significant interest to statisticians and polling experts due to its potential impact on polling models and election forecasting.
+To provide some background, President Biden ran on his merit as a former Vice President of eight years and had faced lots of criticism during his time in office.
+Moreover, Former President Trump had controversial and career-damaging events.
+During the 2016 elections, Trump even faced Before the 2020 elections, Trump was charged with misconduct while in office(cite).
+However, the senate acquitted him of those charges (cite).
+Thus, both Biden and Trump faced criticism from the electorate.
+An interesting question to consider is despite these, what characteristics did the voters have?
+Do they tell us a story on the type of people that voted for Biden and the type that voted for Trump?
+Answering these questions can help develop polling models and future election forecasting.
+
+For our paper, we used a logistic regression model to estimate the likelihood of voting for Biden or Trump given certain characteristics.
+We use a binary outcome variable and three predictor variables.
+Our estimand is the probability that an individual voted for Biden or Trump based on three characteristics, their gender, race, and ownership of a gun.
+The reason for this choice will be further examined in @sec-model.
+This data driven approach to elections can bring insight to voter characteristics and the importance of survey data and knowing the electorate.
+
+[insert more of what was done and what was found - overview of results section]
+
+The survey data for this paper is from the Harvard Dataverse Repository.
+The Cooperative Election Study is a sample survey that consists of pre-election and post-election questions.
+They provide a guide and a dataset for numerous years.
+It is also open to the public.
+[add more]
+
+Data for this analysis and the different applications used will be further introduced in @sec-data.
+@sec-model provides the model set-up and the justification for the use of that model.
+@sec-results will show results and @sec-discussion is a discussion of the results including the paper’s weaknesses and biases.
+
 # Data {#sec-data}
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2024-03-12 paper/paper.qmd
Add information section to paper

@@
 Moreover, Former President Trump had controversial and career-damaging events.
-During the 2016 elections, Trump even faced Before the 2020 elections, Trump was charged with misconduct while in office(cite).
-However, the senate acquitted him of those charges (cite).
+During the 2016 lections, Trump even faced Before the 2020 elections, Trump was charged with misconduct while in office(cite).
+However, the senate acquitted him of those charges.
 Thus, both Biden and Trump faced criticism from the electorate.

### 2024-03-12 paper/paper.qmd
Add draft of data section to paper

@@
 # Data {#sec-data}
+
+## Sources and Collection
+
+For this paper, we utilized the data from the 2020 Presidential Elections.
+We selected, gender, race, and whether an individual or a family member in the household owns a gun.
+We also selected the variable that recorded whether an individual voted for Biden or Trump.
+The broader context of the dataset is the political landscape of the US during the election cycle of 2020.
+The outcome of the 2020 election and every presidential election usually changes the direction of the country and demonstrates the perspectives of the majority of the electorate.
+There may have been similar dataset available, but we chose this one due to relevance to our research question.
+The questionnaire asked various questions on race, political party preference, voting preference, ownership of a gun and have numerous multiple choice answers available.
+Thus, this survey data provided us with the information on the demographic characteristics we wanted to consider.
+
+Our paper aims to address the following two questions: (1) Does gender and race play a role in political preference? (2) Do those that own a gun favour one candidate over the other?
+
+## Methodology
+
+## Variables
+
+## Measurement
 

### 2024-03-12 paper/paper.qmd
added graphs

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Voter Characteristics and Political Support in the US"
+# subtitle: "My subtitle if needed"
 author: 
@@
   - Rahma Binth Mohammad
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/atn-ly/political_support_in_the_us"
 date: today
@@
 library(tidyverse)
+library(knitr)
+library(here)
+library(ggplot2)
+library(arrow)
+library(scales)
+
+#### Read in cleaned data ####
+
+ces2020 = read_parquet(
+  file = here("data/analysis_data/ces2020_analysis_data.parquet"),
+  show_col_types = FALSE
+)
 ```
@@
 
-# Results {#sec-results}
+# Results [In progress...] {#sec-results}
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tmp-label1
+#| tbl-cap: tmp-caption1
+
+ces2020 |>
+  ggplot(aes(x = race, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender)) +
+  theme_minimal() +
+  labs(
+    x = "Race",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_x_discrete(limits = rev) +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tmp-label2
+#| tbl-cap: tmp-caption2
+
+ces2020 |>
+  ggplot(aes(x = gun_ownership, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender)) +
+  theme_minimal() +
+  labs(
+    x = "Gun Ownership",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_x_discrete(limits = rev, labels = label_wrap(10)) +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+
+```
 

### 2024-03-12 scripts/02-data_cleaning.R

@@ -63,18 +63,20 @@ ces2020 <-
       )
     ),
     gun_ownership = case_when(
-      gunown == 1 ~ "Personally own a gun",
-      gunown == 2 ~ "Don't personally own a gun, but someone in the household owns a gun",
-      gunown == 3 ~ "No one in the household owns a gun",
-      gunown == 4 ~ "Not sure",
+      gunown == 1 ~ "Personal",
+      gunown == 2 ~ "Someone in the household",
+      gunown == 3 ~ "No one in the household",
+      gunown == 8 ~ "Not sure",
+      TRUE ~ "No Answer"  # NA values, set to No Answer
     ),
     gun_ownership = factor(
       gun_ownership,
       levels = c(
-        "Personally own a gun",
-        "Don't personally own a gun, but someone in the household owns a gun",
-        "No one in the household owns a gun",
-        "Not sure"
+        "Personal",
+        "Someone in the household",
+        "No one in the household",
+        "Not sure",
+        "No Answer"
       )
     ),
   ) |>
@@ -83,4 +85,4 @@ ces2020 <-
 ces2020
 
 #### Save data ####
-write_parquet(ces2020, "data/analysis_data/analysis_data.parquet")
+write_parquet(ces2020, "data/analysis_data/ces2020_analysis_data.parquet")

### 2024-03-12 paper/paper.qmd
Add paragraphs to model justification section

@@
 ## Model justification
+[Paragraph 1: How features enter the model and why → link with data section]
+
+We chose gender, race, and gun ownership status to be the predictor variables in our model.
+These features were chosen based on our data section.
+[more explanation] First, we expect a negative relationship between men and support for Biden.
+[more explanation] Second, we expect a positive relationship between non-white respondents and support for Biden.
+[more explanation] Third, we expect a negative relationship between gun owners and support for Biden.
+[more explanation]
+
+[Paragraph 2: Assumptions that underpin the model]
+
+Logistic regression does not make the same assumptions as linear regression.
+First, linear regression assumes a continuous outcome variable that can take any number on the real line, whereas logistic regression assumes a binary outcome variable.
+Furthermore, linear regression requires that the outcome variable is a linear function of the predictor variables, while the outcome in logistic regression is part of the exponential family.
+However, logistic regression does assume that the relationship between the log-odds of the binary outcome and predictor variables is linear.
+Lastly, unlike in linear regression, logistic regression does not require the assumption of homoscedasticity of errors.
+
+[Paragraph 3: Alternative models or variants; Strengths and weaknesses; Circumstances in which the model may not be appropriate]
+
+Alternative regression models were considered, but rejected because they were not appropriate for our outcome variable.
+Linear models assume a continuous outcome variable, and Poisson and Negative binomial models assume count outcome variables.
+Since our outcome variable is binary, we chose to use logistic regression.
+
+[Paragraph 4: Evidence of model validation and checking, model convergence, and/or diagnostic issues]
 

### 2024-03-14 scripts/04-model.R
Change model to n=1000 samples

@@ -18,10 +18,15 @@ ces2020_analysis_data <- read_parquet("data/analysis_data/ces2020_analysis_data.
 ### Model data ####
 set.seed(321)
 
+# Randomly sample 1000 observations in the interest of run-time
+ces2020_reduced <- 
+  ces2020_analysis_data |> 
+  slice_sample(n = 1000)
+
 political_preferences <-
   stan_glm(
     voted_for ~ gender + race + gun_ownership,
-    data = ces2020_analysis_data,
+    data = ces2020_reduced,
     family = binomial(link = "logit"),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-03-14 scripts/04-model.R
Update model with n=5000 samples

@@ -18,10 +18,10 @@ ces2020_analysis_data <- read_parquet("data/analysis_data/ces2020_analysis_data.
 ### Model data ####
 set.seed(321)
 
-# Randomly sample 1000 observations in the interest of run-time
+# Randomly sample 5000 observations in the interest of run-time
 ces2020_reduced <- 
   ces2020_analysis_data |> 
-  slice_sample(n = 1000)
+  slice_sample(n = 5000)
 
 political_preferences <-
   stan_glm(

### 2024-03-14 paper/paper.qmd
Change model to n=1000 samples
Added finished model section

@@
 
+#### Workspace setup ####
+
 library(tidyverse)
+library(rstanarm)
 library(knitr)
@@
 library(scales)
+library(modelsummary)
 
-#### Read in cleaned data ####
-
+# Read in cleaned data
 ces2020 = read_parquet(
@@
 )
+
+# Read in model
+political_preferences <-
+  readRDS(file = here::here("models/political_preferences.rds"))
+
 ```
@@
 
-The goal of our modeling strategy is to forecast if a person voted for Biden, based only on knowing their gender, race, and gun ownership status.
+The goal of our modeling strategy is to forecast if a person is likely to vote for Biden, based only on knowing their gender, race, and gun ownership status.
 
@@
 
-We used a logistic regression model in a Bayesian framework using the package `rstanarm` [@rstanarm], which we will briefly describe here.
+We used a logistic regression model in a Bayesian framework using the package `rstanarm` [@rstanarm] in R [@citeR], which we will briefly describe here.
 Logistic regression is a type of generalized linear model.
@@
 The parameter $\alpha$ is the intercept and the parameters, $\beta$, $\gamma$, and $\delta$, are the slope coefficients.
-We specify prior probability distributions for each of the parameters in our model, but these are just the default priors that `rstanarm` [@rstanarm] uses (Normal distribution with mean and standard deviation of 0 and 2.5).
+We specify prior probability distributions for each of the parameters in our model.
+However, these are just the default priors that `rstanarm` [@rstanarm] uses, which are Normal distributions with a mean and standard deviation of 0 and 2.5, respectively.
 
 ## Model justification
-[Paragraph 1: How features enter the model and why → link with data section]
 
-We chose gender, race, and gun ownership status to be the predictor variables in our model.
-These features were chosen based on our data section.
-[more explanation] First, we expect a negative relationship between men and support for Biden.
-[more explanation] Second, we expect a positive relationship between non-white respondents and support for Biden.
-[more explanation] Third, we expect a negative relationship between gun owners and support for Biden.
-[more explanation]
-
-[Paragraph 2: Assumptions that underpin the model]
+Given that Biden and Trump are far apart on women’s issues, racial justice, and gun policy, we  chose *gender*, *race*, and *gunown* to be the predictor variables in our model.
+First, since we are interested in forecasting who a respondent is likely to vote for, we used *gender* instead of *gender_post* because it is representative of all adult Americans rather than only adult Americans who are registered to vote.
+Second, we used *race* instead of variables by country or region to affirm that race is a social construct with no biological foundation.
+Third, we used *gunown* instead of variables about gun control because it captures broader opinions rather than specific policy proposals.
+After exploring the data in the previous section, we found differences in political preference based on these features and that it would be of interest to investigate further.
 
@@
 
-[Paragraph 3: Alternative models or variants; Strengths and weaknesses; Circumstances in which the model may not be appropriate]
-
 Alternative regression models were considered, but rejected because they were not appropriate for our outcome variable.
@@
 
-[Paragraph 4: Evidence of model validation and checking, model convergence, and/or diagnostic issues]
+To show that our model does a good job of fitting the data, we consider the posterior distribution and implement posterior predictive checks.
+Details and graphs can be found in @sec-model-details.
+Furthermore, we check if the Markov Chain Monte Carlo (MCMC) sampling algorithm that `rstanarm` uses to obtain samples from the posterior distributions of interest ran into any issues.
+Details and graphs can be found in @sec-model-details.
 
@@
 
+## Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
+It shows a very close match between the actual outcome variable with simulations from the posterior distribution.
+This suggests that our model does a good job of fitting the data.
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
+It shows that estimates changed minimally once the data was taken into account.
+This suggests that we specified good priors.
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+# Read data
+ces2020_analysis_data <- read_parquet(here::here("data/analysis_data/ces2020_analysis_data.parquet"))
+
+set.seed(321)
+
+# Randomly sample 5000 observations in the interest of run-time
+ces2020_reduced <- 
+  ces2020_analysis_data |> 
+  slice_sample(n = 5000)
+
+pp_check(political_preferences) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(political_preferences) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
+
+## Diagnostics
+
+@fig-diagnostics-1 is a trace plot.
+It shows that there are no horizontal lines that appear to bounce around and have a nice overlap between the chains.
+This does not suggest anything out of the ordinary.
+
+@fig-diagnostics-2 is a Rhat plot. It shows that everything is close to 1 and no more than 1.1. This does not suggest any problems.
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-diagnostics
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
+
+plot(political_preferences, "trace")
+
+plot(political_preferences, "rhat")
+```
+
 \newpage

### 2024-03-16 paper/paper.qmd
Added Results section and changed data_cleaning.R

@@
 library(modelsummary)
+library(marginaleffects )
 
@@
 # Data {#sec-data}
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tmp-label1
+#| tbl-cap: tmp-caption1
+
+ces2020 |>
+  ggplot(aes(x = race, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender)) +
+  theme_minimal() +
+  labs(
+    x = "Race",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_x_discrete(limits = rev) +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tmp-label2
+#| tbl-cap: tmp-caption2
+
+ces2020 |>
+  ggplot(aes(x = gun_ownership, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(gender)) +
+  theme_minimal() +
+  labs(
+    x = "Gun Ownership",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_x_discrete(limits = rev, labels = label_wrap(10)) +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+
+```
 
@@
 
-# Results [In progress...] {#sec-results}
+# Results {#sec-results}
 
@@
 #| message: false
-#| label: tmp-label1
-#| tbl-cap: tmp-caption1
+#| warning: false
+#| label: tbl-model-summary-table
+#| tbl-cap: "Explanatory models of political preferences based on gender, race, and gun ownership (n = 5000)"
+library(modelsummary)
 
-ces2020 |>
-  ggplot(aes(x = race, fill = voted_for)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(gender)) +
-  theme_minimal() +
-  labs(
-    x = "Race",
-    y = "Number of respondents",
-    fill = "Voted for"
-  ) +
-  coord_flip() +
-  scale_x_discrete(limits = rev) +
-  scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
+modelsummary(
+  list(
+    "Support Biden" = political_preferences
+  ),
+  statistic = "mad"
+  )
+```
 
-```
+We performed logistic regression analysis on 5000 observations; a subset of the total 43240 observations from our cleaned dataset.
+Since we are interested in which presidential candidate an individual voted for based on the categories explained earlier, our model made `genderMale`, `raceWhite`, `gun_ownershipPersonal` into reference groups for their respective categories.
+
+Our results from the model given by @tbl-model-summary-table shows the coefficient of each indicator variable.
+This coefficient represents the log of the expected difference in support for Biden compared to Trump.
+When this value is negative, it shows a decrease in support for Biden, when this value is 0 it shows neither a decrease nor increase in support for Biden.
+When this value is positive, it shows an increase in support for Biden.
+
+The intercept is -0.748 which means that the support for Biden decreases by about 0.784 units when all other variables are held constant.
+By holding all variables constant, we have an individual who is Male, White, and personally owns a gun.
+
+## Gender
+
+We can see that for the variable `genderFemale`, we have a coefficient of 0.235 indicating an increase in support for Biden over the reference group being `genderMale`, while keeping the other variables constant.
+
+## Race
+
+We can see that Black individuals show a much stronger support for Biden as the coefficient for `raceBlack` is 2.464.
+Asian, Hispanic, and Middle Eastern individuals all show an increase in support for Biden with coefficients of 0.790, 0.548, and 0.288 respectively.
+Native Americans show a decrease in support for Biden with a coefficient of -0.536 while individuals of two or more races show neither an increase nor decrease in support for Biden with a coefficient of -0.008.
+
+## Gun Ownership
+
+For gun ownership, we can see that if an individual does not personally own a gun but someone in their household does, the coefficient is 0.870 which shows an increase in support for Biden.
+If nobody in an individual’s household owns a gun, the coefficient is 1.332 showing a strong support for Biden.
+When an individual does not own a gun and is not sure if anyone in their household owns a gun, the coefficient is -0.815 showing a decrease in support for Biden.
 
@@
 #| message: false
-#| label: tmp-label2
-#| tbl-cap: tmp-caption2
+#| warning: false
+#| label: tbl-biden-or-trump-predictions
+#| tbl-cap: "Probability that an individual supports Biden given their demographic"
+political_predictions <-
+  predictions(political_preferences)
 
-ces2020 |>
-  ggplot(aes(x = gun_ownership, fill = voted_for)) +
-  stat_count(position = "dodge") +
-  facet_wrap(facets = vars(gender)) +
-  theme_minimal() +
+knitr::kable(head(political_predictions))
+```
+
+From the model, we used `predictions()` from `marginaleffects` [@marginaleffects] to obtain the implied probability that an individual supports Biden.
+A sample of the result is given in @tbl-biden-or-trump-predictions where each individual is given an `estimate` value which is the estimated probability that the individual supports Biden.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-model-gender
+#| fig-cap: "Predicted support of Biden given gender vs. actual support"
+political_predictions |>
+  mutate(gender = factor(gender)) |>
+  ggplot(aes(x = gender, y = estimate, color = voted_for)) +
+  geom_jitter(width = 0.01, height = 0.01, alpha = 0.3) +
+  labs(
+    x = "Gender",
+    y = "Estimated probability that an \n individual voted for Biden",
+    color = "Candidate actually voted for"
+  ) +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+```
+
+Using the results in @tbl-biden-or-trump-predictions, we plotted three graphs using `ggplot2` [@ggplot2] which shows the estimated probability that an individual supports Biden given gender, race, and gun ownership with the colour of the point indicating their true support.
+@fig-model-gender shows the estimated probability that an individual supports Biden given gender.
+To support the accuracy of our model, blue dots should be higher up on the graph while red dots should be lower down indicating that our predicted probability matches with the actual result.
+We can see in @fig-model-gender that this trend holds true.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-model-race
+#| fig-cap: "Predicted support of Biden given race vs. actual support"
+political_predictions |>
+  mutate(race = factor(race)) |>
+  ggplot(aes(x = race, y = estimate, color = voted_for)) +
+  geom_jitter(width = 0.01, height = 0.01, alpha = 0.3) +
+  labs(
+    x = "Race",
+    y = "Estimated probability that an \n individual voted for Biden",
+    color = "Candidate actually voted for"
+  ) +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  scale_x_discrete(guide = guide_axis(n.dodge=2))
+```
+
+@fig-model-race shows the estimated probability that an individual supports Biden given race.
+There appears to be red dots near the top for White, Native American, Middle Eastern, and Two or more races which indicates possible inaccuracy in our model.
+However, Black, Hispanic and Asian appears to follow the expected distribution of the dots.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-model-gun-owner
+#| fig-cap: "Predicted support of Biden given gun ownership vs. actual support"
+political_predictions |>
+  mutate(gun_ownership = factor(gun_ownership)) |>
+  ggplot(aes(x = gun_ownership, y = estimate, color = voted_for)) +
+  geom_jitter(width = 0.01, height = 0.01, alpha = 0.3) +
   labs(
     x = "Gun Ownership",
-    y = "Number of respondents",
-    fill = "Voted for"
+    y = "Estimated probability that an \n individual voted for Biden",
+    color = "Candidate actually voted for"
   ) +
-  coord_flip() +
-  scale_x_discrete(limits = rev, labels = label_wrap(10)) +
-  scale_fill_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
+  theme_classic() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  scale_x_discrete(guide = guide_axis(n.dodge=2))
+```
 
-```
+@fig-model-gun-owner shows the estimated probability that an individual supports Biden given their gun ownership status.
+We can see that Personal gun ownership and the Not Sure if anyone in the household owns a gun categories show a lower support of Biden, while the No One in the household and the Someone else in the household owns a gun categories show a higher trend of support for Biden.
+There appears to be an expected cluster of blue dots higher up and red dots lower down, which supports the accuracy of our model.
 

### 2024-03-16 scripts/02-data_cleaning.R

@@ -32,7 +32,7 @@ ces2020 <-
 
 ces2020 <-
   ces2020 |>
-  filter(votereg == 1,
+  filter(!is.na(gunown), votereg == 1,
          CC20_410 %in% c(1, 2)) |>
   mutate(
     voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),

### 2024-03-16 scripts/00-simulate_data.R
Added tests script

@@ -30,3 +30,34 @@ us_political_preferences <- tibble(
   select(-support_prob, supports_trump, race, gender, gun_ownership)
 
 us_political_preferences
+
+#### Test simulated data ####
+# Check number of observations
+nrow(us_political_preferences) == 1000
+# Check correct type gender
+all(sapply(us_political_preferences$gender, is.character))
+# Check Only two values in us_political_preferences$gender
+length(unique(us_political_preferences$gender)) == 2
+# Check correct values in us_political_preferences$gender
+all(c("Male", "Female") %in% us_political_preferences$gender)
+
+# Check correct type race
+all(sapply(us_political_preferences$race, is.character))
+# Check Only 2 values in us_political_preferences$race
+length(unique(us_political_preferences$gender)) == 2
+# Check correct values in us_political_preferences$race
+all(c("White", "Not White") %in% us_political_preferences$race)
+
+# Check correct type gun_ownership
+all(sapply(us_political_preferences$gun_ownership, is.character))
+# Check Only 2 values in us_political_preferences$gun_ownership
+length(unique(us_political_preferences$gun_ownership)) == 2
+# Check correct values in us_political_preferences$gun_ownership
+all(c("Owns a gun", "Doesn't own a gun") %in% us_political_preferences$gun_ownership)
+
+# Check correct type supports_trump
+all(sapply(us_political_preferences$supports_trump, is.character))
+# Check Only 2 values in us_political_preferences$supports_trump
+length(unique(us_political_preferences$supports_trump)) == 2
+# Check correct values in us_political_preferences$supports_trump
+all(c("yes", "no") %in% us_political_preferences$supports_trump)

### 2024-03-16 scripts/02-data_cleaning.R

@@ -67,7 +67,6 @@ ces2020 <-
       gunown == 2 ~ "Someone in the household",
       gunown == 3 ~ "No one in the household",
       gunown == 8 ~ "Not sure",
-      TRUE ~ "No Answer"  # NA values, set to No Answer
     ),
     gun_ownership = factor(
       gun_ownership,
@@ -75,8 +74,7 @@ ces2020 <-
         "Personal",
         "Someone in the household",
         "No one in the household",
-        "Not sure",
-        "No Answer"
+        "Not sure"
       )
     ),
   ) |>

### 2024-03-16 scripts/03-test_data.R

@@ -1,15 +1,50 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests for ces2020_analysis_data.parquet file
+# Author: Renfrew Ao-Ieong, Rahma Binth Mohammad, Tam Ly
+# Date: 15 March 2024
+# Contact: renfrew.aoieong@mail.utoronto.ca, rahma.binthmohammad@mail.utoronto.ca, annatn.ly@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Run the "02-data.cleaning.R" script to generate ces2020_analysis_data.parquet
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(arrow)
+
+#### Read data ####
+ces2020_analysis_data <- read_parquet("data/analysis_data/ces2020_analysis_data.parquet")
 
 #### Test data ####
+# Check dataset contains 43240 observations (for reproducibility)
+nrow(ces2020_analysis_data) == 43240
+# Check voted_for is a factor and includes Biden and Trump only
+all(sapply(ces2020_analysis_data$voted_for, is.factor))
+length(levels(ces2020_analysis_data$voted_for)) == 2
+"Trump" %in% levels(ces2020_analysis_data$voted_for) && "Biden" %in% levels(ces2020_analysis_data$voted_for)
+
+# Check gender is a factor only including 1 and 2 (Male, Female respectively)
+all(sapply(ces2020_analysis_data$gender, is.factor))
+length(levels(ces2020_analysis_data$gender)) == 2
+"Male" %in% levels(ces2020_analysis_data$gender) && "Female" %in% levels(ces2020_analysis_data$gender)
+
+# Check race is an factor [1-8] and all races in dataset are included
+all(sapply(ces2020_analysis_data$race, is.factor))
+length(levels(ces2020_analysis_data$race)) == 8
+
+"White" %in% levels(ces2020_analysis_data$race) && 
+"Black" %in% levels(ces2020_analysis_data$race) &&
+"Hispanic" %in% levels(ces2020_analysis_data$race) &&
[…]

### 2024-03-16 paper/paper.qmd
Inline citation for code

@@
 #| tbl-cap: tmp-caption1
-
+# Code adapted from https://tellingstorieswithdata.com/13-ijaglm.html#logistic-regression
 ces2020 |>
@@
 #| tbl-cap: tmp-caption2
-
+# Code adapted from https://tellingstorieswithdata.com/13-ijaglm.html#logistic-regression
 ces2020 |>
@@
 #| fig-cap: "Predicted support of Biden given gender vs. actual support"
+
+# Code adapted from https://tellingstorieswithdata.com/13-ijaglm.html#logistic-regression
 political_predictions |>
@@
 #| fig-cap: "Predicted support of Biden given race vs. actual support"
+
+# Code adapted from https://tellingstorieswithdata.com/13-ijaglm.html#logistic-regression
 political_predictions |>
@@
 #| fig-cap: "Predicted support of Biden given gun ownership vs. actual support"
+
+# Code adapted from https://tellingstorieswithdata.com/13-ijaglm.html#logistic-regression
 political_predictions |>

### 2024-03-16 paper/paper.qmd
Add abstract and discussion section

@@
 ---
-title: "Voter Characteristics and Political Support in the US"
+title: "Gender, race, and gun policy during the 2020 US Presidential Election: Biden more favorable among women, Black, Asian, Hispanic, Middle Eastern, and non-gun owners"
 # subtitle: "My subtitle if needed"
@@
   - Rahma Binth Mohammad
-thanks: "Code and data are available at: https://github.com/atn-ly/political_support_in_the_us"
+thanks: "Code and data are available at: [https://github.com/atn-ly/political_support_in_the_us](https://github.com/atn-ly/political_support_in_the_us)"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "The outcome of the 2020 US Presidential Election served as a reflection of the beliefs and attitudes held by adult Americans.
+Using data obtained during the fall of 2020 from the Cooperative Election Study, we investigated if a person is likely to vote for Biden, based only on knowing their gender, race, and gun ownership status.
+The results of our analysis using a Bayesian logistic regression model show that women, Black, Asian, Hispanic, Middle Eastern, and non-gun owners are more likely to support Biden.
+Our discovery highlights the importance of women’s issues, racial justice, and gun policy on American voters."
 format: pdf
@@
 bibliography: references.bib
+toc: true
 ---
@@
 library(modelsummary)
-library(marginaleffects )
+library(marginaleffects)
 
@@
 
-Given that Biden and Trump are far apart on women’s issues, racial justice, and gun policy, we  chose *gender*, *race*, and *gunown* to be the predictor variables in our model.
-First, since we are interested in forecasting who a respondent is likely to vote for, we used *gender* instead of *gender_post* because it is representative of all adult Americans rather than only adult Americans who are registered to vote.
-Second, we used *race* instead of variables by country or region to affirm that race is a social construct with no biological foundation.
-Third, we used *gunown* instead of variables about gun control because it captures broader opinions rather than specific policy proposals.
+Given that Biden and Trump are far apart on women’s issues, racial justice, and gun policy, we  chose “gender”, “race”, and “gunown” to be the predictor variables in our model.
+First, since we are interested in forecasting who a respondent is likely to vote for, we used “gender” instead of “gender_post” because it is representative of all adult Americans registered to vote rather than only adult Americans who voted.
+Second, we used “race” instead of variables by country or region to affirm that race is a social construct with no biological foundation.
+Third, we used “gunown” instead of variables about gun control because it captures broader opinions rather than specific policy proposals.
 After exploring the data in the previous section, we found differences in political preference based on these features and that it would be of interest to investigate further.
@@
 To show that our model does a good job of fitting the data, we consider the posterior distribution and implement posterior predictive checks.
-Details and graphs can be found in @sec-model-details.
 Furthermore, we check if the Markov Chain Monte Carlo (MCMC) sampling algorithm that `rstanarm` uses to obtain samples from the posterior distributions of interest ran into any issues.
@@
 
-## First discussion point {#sec-first-point}
+## Findings
 
-## Second discussion point
+In this paper, data from the 2020 Cooperative Election Study (CES) [@CES] of US political opinion was downloaded and cleaned as discussed in @sec-data.
+Then, we used a Bayesian logistic regression model as seen in @sec-model to forecast if a person is likely to support Biden, based only on knowing their gender, race, and gun ownership status.
+Following the outcomes of our analysis in @sec-results, we summarize our three major results here:
 
-## Third discussion point
+1. Women are more likely to vote for Biden,
+2. Black, Asian, Hispanic, and Middle Eastern individuals are more likely to vote for Biden, and
+3. Non-gun owners are more likely to vote for Biden.
+
+We will now justify and detail the implications of our results by providing the political, historical, cultural, and social context that shaped our story of American political preference in the 2020 US presidential election.
+
+## Women are more likely to vote for Biden
+
+In @sec-results, we found that women have a coefficient of 0.235 and higher estimated probabilities that our model implies of supporting Biden.
+This means that by just knowing that an individual is female, we can forecast that they are more likely to vote for Biden.
+Our results imply that Biden’s policies on women’s issues are more favorable than Trump’s.
+This confirms what we expected as Biden and Trump held opposing stances on various important women’s issues during the 2020 election including women’s reproductive rights, women’s health, and violence against women.
+
+Abortion is a key issue and a deal-breaker for many voters during an election.
+During Biden’s 2020 campaign, he vowed to protect a women’s right to choose and keep access to abortion legal [@repro_rights].
+In particular, Biden promised to protect and codify *Roe v Wade*, a landmark Supreme Court ruling from 1973, which decriminalized abortion nationwide [@repro_rights].
+In contrast, Trump pushed for the overturning of *Roe* and appointed conservative anti-abortion judges during his presidency [@repro_rights].
+
+Moreover, women’s health is another important issue that voters consider when deciding which candidate to support.
+During the 2020 election, Biden supported women’s access to coverage and care [@health].
+For example, Biden campaigned to protect the 2010 *Affordable Care Act*, which provides affordable healthcare to millions of women in the US and restricts discriminatory insurance policies [@health].
+This contrasts with Trump’s stance on women’s health in which he supported to overturn the entire *ACA* with no plan to replace it [@health].
+
+Furthermore, the issue of sexual violence has been of increasing importance as 1 in 3 women and 1 in 4 men in the US report experiencing sexual violence in their lifetimes [@health].
+During his time as a senator, Biden passed the 1994 *Violence Against Women’s Act* which helped establish many violence prevention efforts including rape crisis centers, shelters, and other support services [@health].
+During his 2020 campaign, Biden supported reauthorizing and expanding this law, whereas Trump allowed the law to expire in 2018 during his presidency and stated no position on his campaign website [@health].
+
+We learn that Biden and Trump have stark differences on women’s issues and that the outcome of the 2020 election would have major consequences for women.
+As seen in our findings, women are more likely to vote for Biden, reflecting their support for policies that protect their rights.
+
+## Black, Asian, Hispanic, and Middle Eastern individuals are more likely to vote for Biden
+
+In @sec-results, we found that Black, Asian, Hispanic, and Middle Eastern individuals have coefficients of 2.464, 0.790, 0.548, and 0.288 respectively and higher estimated probabilities that our model implies of supporting Biden.
+This means that by just knowing that an individual is Black, Asian, Hispanic, or Middle Eastern, we can forecast that they are more likely to vote for Biden.
+Our results imply that Biden’s policies on racial justice are more favorable within marginalized communities than Trump’s.
+This aligns with our expectations as Biden and Trump held opposing stances on various important racial issues during the 2020 election including Black Lives Matter, minority-owned small businesses, and housing and education.
+
+In 2020, the police killings of George Floyd in May and Breonna Taylor in March sparked the largest racial justice protests across the US [@BLM].
+As a result, racial equality was at the center of many voter’s minds during the 2020 election.
+Biden’s legacy of Black support was demonstrated by his choice of Kamala Harris as his running mate, as she is the first woman, the first Black American, and the first South Asian American to hold this position [@BLM].
+In contrast, Trump called Black Lives Matter “a symbol of hate” and encouraged police violence against protestors [@BLM].
+
+Additionally, Biden and Trump had different strategies to address racial disparities with the US including racial wealth, opportunity, and job gaps [@equality].
+In 2020, Biden pledged to support minority-owned small businesses with a 26-page plan by investing $30 billion toward a Small Business Opportunity Fund, $50 billion in venture capital for Black and brown entrepreneurs, and $100 billion toward low-interest loans for new entrepreneurs [@equality].
+This is in stark contrast to Trump, who had no broad policy plan to address racial economic inequality [@equality].
+
+Housing and education is another important issue in combating racial inequality in the US.
+Biden proposed to boost homeownership among minority communities by creating a housing-plan that includes constructing 1.5 million homes and public housing units and ending discriminatory housing policies [@equality].
+On education reform, Biden proposed to make public universities and private historically Black colleges and universities tuition-free for families with incomes under $125,000—which the campaign estimates to apply to about 90% of Black, Latino, and Native American households [@equality].
+On the other hand, Trump had no plans in place to address these issues again [@equality].
+
+We learn that Biden and Trump are far apart on racial justice and that the 2020 election would shape the lives of many non-white people for years to come .
+As seen in our findings, Black, Asian, Hispanic, and Middle Eastern individuals are more likely to vote for Biden, reflecting their support for anti-racist policies that promote equality.
+
+## Non-gun owners are more likely to vote for Biden
+
+In @sec-results, we found that an individual with a non-gun owning household has a coefficient of 1.332 and higher estimated probabilities that our model implies of supporting Biden.
+This means that by just knowing that an individual whose household does not own a gun, we can forecast that they are more likely to vote for Biden.
+Our results imply that Biden’s policies on gun policies for non-gun owners were more favorable than Trump’s.
+This confirms what we expected as Biden and Trump held opposing stances on various important policies including gun control, universal background checks, and assault weapons bans.
+
+With the rise in mass shootings and gun-related deaths across the country, gun policies are as important as ever [@guns].
+Under the Trump administration, the US saw record-breaking tragedies in the 2017 Las Vegas Strip shooting, 2017 Sutherland Springs church shooting, 2018 Parkland high school shooting, and 2019 El Paso shooting [@guns].
+During his presidency, Trump supported the exercise of Second Amendment rights and pushed for greater freedom for gun-owners [@guns2].
+He loosened gun regulations on firearms exports and reversed a law that allowed people deemed mentally unfit to purchase guns [@guns2].
+In contrast, in Biden’s 2020 campaign, he included plans to require universal background checks for all gun sales and bans on semiautomatic rifles and magazines that contained more than 10 bullets [@guns].
+
+We learn that Biden and Trump have opposite approaches to gun policies and that the outcome of the 2020 election would decide the safety of millions of Americans.
+As seen in our findings, non-gun owners are more likely to vote for Biden, reflecting their support for policies that prevent gun violence.
 
 ## Weaknesses and next steps
+
+Weaknesses arise in the data and model that was used.
+Although the sampling methodology detailed in @CES is an accepted approach that balances sampling concerns and cost, there are still limitations with large survey data that must be addressed.
+First, while the survey was not directly conducted by the government, it was conducted by YouGov, an international online research data and analytics technology group, that was funded by the government [@yougov].
+This affects our findings since their clients could be Democrats or Republicans that could have their own partisan goals in mind [@yougov].
+Historically, there have been cases of data manipulation to suit a government’s narrative in the past [@rohan].
+Moreover, there will always be missing data no matter how good the data acquisition process is [@rohan].
+CES respondents are people who have made an account on yougov.com or recruited live from online advertisements, and the 2020 CES survey was conducted over the internet [@CES].
+However, 1 in 5 US households are not connected to the internet at home [@ntia].
+This affects our findings since the sample is representative of Americans with internet access rather than of all national adults as @CES claims.
+
+In future work, we could include additional predictor variables to get a more complete picture of the opinions and traits of voters.
+For example, the results using family income, educational level, and state of residence as variables, along with their implications, could be of interest.
+Moreover, we could include the other two presidential candidates of the Libertarian and Green Party in future studies, which requires a different model.
+Lastly, another aspect to consider could be investigating the post-election questionnaire data.
+We could build the model again using the post-election data and compare the results with our findings from the pre-election data.
 

### 2024-03-16 scripts/03-test_data.R
Add more tests

@@ -1,3 +1,4 @@
+
 #### Preamble ####
 # Purpose: Tests for ces2020_analysis_data.parquet file
 # Author: Renfrew Ao-Ieong, Rahma Binth Mohammad, Tam Ly
@@ -13,6 +14,10 @@ library(arrow)
 
 #### Read data ####
 ces2020_analysis_data <- read_parquet("data/analysis_data/ces2020_analysis_data.parquet")
+
+#### Read model ####
+political_preferences <-
+  readRDS(file = here::here("models/political_preferences.rds"))
 
 #### Test data ####
 # Check dataset contains 43240 observations (for reproducibility)
@@ -32,19 +37,47 @@ all(sapply(ces2020_analysis_data$race, is.factor))
 length(levels(ces2020_analysis_data$race)) == 8
 
 "White" %in% levels(ces2020_analysis_data$race) && 
-"Black" %in% levels(ces2020_analysis_data$race) &&
-"Hispanic" %in% levels(ces2020_analysis_data$race) &&
-"Asian" %in% levels(ces2020_analysis_data$race) &&
-"Native American" %in% levels(ces2020_analysis_data$race) &&
-"Middle Eastern" %in% levels(ces2020_analysis_data$race) &&
-"Two or more races" %in% levels(ces2020_analysis_data$race) &&
-"Other" %in% levels(ces2020_analysis_data$race)
-  
+  "Black" %in% levels(ces2020_analysis_data$race) &&
+  "Hispanic" %in% levels(ces2020_analysis_data$race) &&
+  "Asian" %in% levels(ces2020_analysis_data$race) &&
+  "Native American" %in% levels(ces2020_analysis_data$race) &&
+  "Middle Eastern" %in% levels(ces2020_analysis_data$race) &&
+  "Two or more races" %in% levels(ces2020_analysis_data$race) &&
+  "Other" %in% levels(ces2020_analysis_data$race)
+
 # Check gun_ownership is an factor [1-4] and all categories in dataset are included
 all(sapply(ces2020_analysis_data$gun_ownership, is.factor))
 length(levels(ces2020_analysis_data$gun_ownership)) == 4
 "Personal" %in% levels(ces2020_analysis_data$gun_ownership) && 
-"Someone in the household" %in% levels(ces2020_analysis_data$gun_ownership) &&
-"No one in the household" %in% levels(ces2020_analysis_data$gun_ownership) &&
[…]

### 2024-03-16 paper/paper.qmd
Remove unused file

@@
 #| message: false
-#| label: tmp-label1
-#| tbl-cap: tmp-caption1
+#| label: fig-race-gender
+#| fig-cap: Support for Trump or Biden based on gender and race
 # Code adapted from https://tellingstorieswithdata.com/13-ijaglm.html#logistic-regression
@@
 #| message: false
-#| label: tmp-label2
-#| tbl-cap: tmp-caption2
+#| label: fig-gun-ownership-gender
+#| fig-cap: Support for Trump or Biden based on gender and gun ownership
 # Code adapted from https://tellingstorieswithdata.com/13-ijaglm.html#logistic-regression
@@
 Asian, Hispanic, and Middle Eastern individuals all show an increase in support for Biden with coefficients of 0.790, 0.548, and 0.288 respectively.
-Native Americans show a decrease in support for Biden with a coefficient of -0.536 while individuals of two or more races show neither an increase nor decrease in support for Biden with a coefficient of -0.008.
+Native Americans and Other show a decrease in support for Biden with a coefficient of -0.536 and -0.333 respectively while individuals of two or more races show neither an increase nor decrease in support for Biden with a coefficient of -0.008.
 

### 2024-03-16 paper/paper.qmd
Fix grammatical errors and add references

@@
 ---
-title: "Gender, race, and gun policy during the 2020 US Presidential Election: Biden more favorable among women, Black, Asian, Hispanic, Middle Eastern, and non-gun owners"
-# subtitle: "My subtitle if needed"
+title: "Biden more favorable among female, Black, Asian, Hispanic, Middle Eastern, and non-gun owning voters in the 2020 US Presidential Election"
 author: 
@@
 date-format: long
-abstract: "The outcome of the 2020 US Presidential Election served as a reflection of the beliefs and attitudes held by adult Americans.
+abstract: "The outcome of the 2020 US Presidential Election served as a reflection of the beliefs and attitudes held by Americans.
 Using data obtained during the fall of 2020 from the Cooperative Election Study, we investigated if a person is likely to vote for Biden, based only on knowing their gender, race, and gun ownership status.
-The results of our analysis using a Bayesian logistic regression model show that women, Black, Asian, Hispanic, Middle Eastern, and non-gun owners are more likely to support Biden.
+The results of our analysis using a Bayesian logistic regression model show that female, Black, Asian, Hispanic, Middle Eastern, and non-gun owning individuals are more likely to support Biden.
 Our discovery highlights the importance of women’s issues, racial justice, and gun policy on American voters."
@@
 This means that by just knowing that an individual is Black, Asian, Hispanic, or Middle Eastern, we can forecast that they are more likely to vote for Biden.
-Our results imply that Biden’s policies on racial justice are more favorable within marginalized communities than Trump’s.
+Our results imply that Biden’s policies on racial justice are more favorable than Trump’s within marginalized communities.
 This aligns with our expectations as Biden and Trump held opposing stances on various important racial issues during the 2020 election including Black Lives Matter, minority-owned small businesses, and housing and education.
@@
 
-Additionally, Biden and Trump had different strategies to address racial disparities with the US including racial wealth, opportunity, and job gaps [@equality].
-In 2020, Biden pledged to support minority-owned small businesses with a 26-page plan by investing $30 billion toward a Small Business Opportunity Fund, $50 billion in venture capital for Black and brown entrepreneurs, and $100 billion toward low-interest loans for new entrepreneurs [@equality].
+Additionally, Biden and Trump had different strategies to address racial disparities within the US including wealth, opportunity, and job gaps [@equality].
+In 2020, Biden pledged to support minority-owned small businesses with a 26-page plan that includes investing $30 billion toward a Small Business Opportunity Fund, $50 billion in venture capital for Black and brown entrepreneurs, and $100 billion toward low-interest loans for new entrepreneurs [@equality].
 This is in stark contrast to Trump, who had no broad policy plan to address racial economic inequality [@equality].
@@
 Biden proposed to boost homeownership among minority communities by creating a housing-plan that includes constructing 1.5 million homes and public housing units and ending discriminatory housing policies [@equality].
-On education reform, Biden proposed to make public universities and private historically Black colleges and universities tuition-free for families with incomes under $125,000—which the campaign estimates to apply to about 90% of Black, Latino, and Native American households [@equality].
-On the other hand, Trump had no plans in place to address these issues again [@equality].
+On education reform, Biden proposed to make public universities and private historically Black colleges and universities tuition-free for families with incomes under $125,000 [@equality].
+Again, Trump had no plans in place to address these issues [@equality].
 
-We learn that Biden and Trump are far apart on racial justice and that the 2020 election would shape the lives of many non-white people for years to come .
+We learn that Biden and Trump are far apart on racial justice and that the 2020 election would shape the lives of many non-white people for years to come.
 As seen in our findings, Black, Asian, Hispanic, and Middle Eastern individuals are more likely to vote for Biden, reflecting their support for anti-racist policies that promote equality.
@@
 
-In @sec-results, we found that an individual with a non-gun owning household has a coefficient of 1.332 and higher estimated probabilities that our model implies of supporting Biden.
-This means that by just knowing that an individual whose household does not own a gun, we can forecast that they are more likely to vote for Biden.
-Our results imply that Biden’s policies on gun policies for non-gun owners were more favorable than Trump’s.
+In @sec-results, we found that non-gun owning individuals with non-gun owning households have a coefficient of 1.332 and higher estimated probabilities that our model implies of supporting Biden.
+This means that by just knowing that an individual and whose household does not own a gun, we can forecast that they are more likely to vote for Biden.
+Our results imply that Biden’s gun policies were more favorable than Trump’s for non-gun owners.
 This confirms what we expected as Biden and Trump held opposing stances on various important policies including gun control, universal background checks, and assault weapons bans.
 
-With the rise in mass shootings and gun-related deaths across the country, gun policies are as important as ever [@guns].
+With the rise in mass shootings and gun-related deaths across the country, safe gun policies are as important as ever [@guns].
 Under the Trump administration, the US saw record-breaking tragedies in the 2017 Las Vegas Strip shooting, 2017 Sutherland Springs church shooting, 2018 Parkland high school shooting, and 2019 El Paso shooting [@guns].
-During his presidency, Trump supported the exercise of Second Amendment rights and pushed for greater freedom for gun-owners [@guns2].
+During his presidency, Trump supported Second Amendment rights and pushed for greater freedom for gun-owners [@guns2].
 He loosened gun regulations on firearms exports and reversed a law that allowed people deemed mentally unfit to purchase guns [@guns2].
-In contrast, in Biden’s 2020 campaign, he included plans to require universal background checks for all gun sales and bans on semiautomatic rifles and magazines that contained more than 10 bullets [@guns].
+In contrast, during Biden’s 2020 campaign, he included plans to require universal background checks for all gun sales and bans on semiautomatic rifles and magazines that contained more than 10 bullets [@guns].
 
@@
 First, while the survey was not directly conducted by the government, it was conducted by YouGov, an international online research data and analytics technology group, that was funded by the government [@yougov].
-This affects our findings since their clients could be Democrats or Republicans that could have their own partisan goals in mind [@yougov].
+This affects our findings since their clients could be Democrats or Republicans that have their own partisan goals in mind [@yougov].
 Historically, there have been cases of data manipulation to suit a government’s narrative in the past [@rohan].
@@
 
-In future work, we could include additional predictor variables to get a more complete picture of the opinions and traits of voters.
+In future work, we could include additional predictor variables to get a more complete picture of the opinions and traits of American voters.
 For example, the results using family income, educational level, and state of residence as variables, along with their implications, could be of interest.
-Moreover, we could include the other two presidential candidates of the Libertarian and Green Party in future studies, which requires a different model.
+Moreover, we could include the other two presidential candidates of the Libertarian and Green Party in future studies, which would require a different model.
 Lastly, another aspect to consider could be investigating the post-election questionnaire data.

### 2024-03-16 paper/paper.qmd
Removed draft introduction

@@
 # Introduction
-
-The United States (US) presidential election of 2020 resulted in President Joe Biden’s victory, making him the 46th President of the US.
-The two top runners were Biden for the Democrats and former President Donald Trump for the Republicans.
-The presidential race between the two was of significant interest to statisticians and polling experts due to its potential impact on polling models and election forecasting.
-To provide some background, President Biden ran on his merit as a former Vice President of eight years and had faced lots of criticism during his time in office.
-Moreover, Former President Trump had controversial and career-damaging events.
-During the 2016 lections, Trump even faced Before the 2020 elections, Trump was charged with misconduct while in office(cite).
-However, the senate acquitted him of those charges.
-Thus, both Biden and Trump faced criticism from the electorate.
-An interesting question to consider is despite these, what characteristics did the voters have?
-Do they tell us a story on the type of people that voted for Biden and the type that voted for Trump?
-Answering these questions can help develop polling models and future election forecasting.
-
-For our paper, we used a logistic regression model to estimate the likelihood of voting for Biden or Trump given certain characteristics.
-We use a binary outcome variable and three predictor variables.
-Our estimand is the probability that an individual voted for Biden or Trump based on three characteristics, their gender, race, and ownership of a gun.
-The reason for this choice will be further examined in @sec-model.
-This data driven approach to elections can bring insight to voter characteristics and the importance of survey data and knowing the electorate.
-
-[insert more of what was done and what was found - overview of results section]
-
-The survey data for this paper is from the Harvard Dataverse Repository.
-The Cooperative Election Study is a sample survey that consists of pre-election and post-election questions.
-They provide a guide and a dataset for numerous years.
-It is also open to the public.
-[add more]
-
-Data for this analysis and the different applications used will be further introduced in @sec-data.
-@sec-model provides the model set-up and the justification for the use of that model.
-@sec-results will show results and @sec-discussion is a discussion of the results including the paper’s weaknesses and biases.
 

### 2024-03-16 paper/paper.qmd
Removed draft data section of paper

@@
 ## Sources and Collection
-
-For this paper, we utilized the data from the 2020 Presidential Elections.
-We selected, gender, race, and whether an individual or a family member in the household owns a gun.
-We also selected the variable that recorded whether an individual voted for Biden or Trump.
-The broader context of the dataset is the political landscape of the US during the election cycle of 2020.
-The outcome of the 2020 election and every presidential election usually changes the direction of the country and demonstrates the perspectives of the majority of the electorate.
-There may have been similar dataset available, but we chose this one due to relevance to our research question.
-The questionnaire asked various questions on race, political party preference, voting preference, ownership of a gun and have numerous multiple choice answers available.
-Thus, this survey data provided us with the information on the demographic characteristics we wanted to consider.
-
-Our paper aims to address the following two questions: (1) Does gender and race play a role in political preference? (2) Do those that own a gun favour one candidate over the other?
 

### 2024-03-16 paper/paper.qmd
Update paper with introduction and data

--- 2024-03-16 paper/paper.qmd
@@
 ---
-title: "Biden more favorable among female, Black, Asian, Hispanic, Middle Eastern, and non-gun owning voters in the 2020 US Presidential Election"
+title: "Biden More Favorable Among Female, Black, Asian, Hispanic, Middle Eastern, and Non-gun Owning Voters in the 2020 US Presidential Election"
 author: 
@@
 Moreover, Former President Trump had controversial and career-damaging events.
-During the 2016 lections, Trump even faced Before the 2020 elections, Trump was charged with misconduct while in office(cite).
-However, the senate acquitted him of those charges.
-Thus, both Biden and Trump faced criticism from the electorate.
+During the 2016 elections, Trump even faced accusations regarding election rigging (cite).
+Before the 2020 elections, Trump was charged with misconduct while in office(cite).
+However, the senate acquitted him of those misconduct charges.
+Thus, both Biden and Trump faced have previously criticism from the electorate.
 An interesting question to consider is despite these, what characteristics did the voters have?
@@
 
-For our paper, we used a logistic regression model to estimate the likelihood of voting for Biden or Trump given certain characteristics.
+For our paper, we used a logistic regression model to estimate the likelihood of individuals voting for Biden or Trump, given certain characteristics.
 We use a binary outcome variable and three predictor variables.
-Our estimand is the probability that an individual voted for Biden or Trump based on three characteristics, their gender, race, and ownership of a gun.
+Our estimand is the probability that an individual voted for Biden or Trump based on three characteristics; their gender, race, and ownership of a gun.
 The reason for this choice will be further examined in @sec-model.
 This data driven approach to elections can bring insight to voter characteristics and the importance of survey data and knowing the electorate.
+We sought to produce results that could be used to predict voter behavior in the future.
 
-[insert more of what was done and what was found - overview of results section]
+Based on our model and analysis, we found that certain demographic groups were mor include to support either candidate.
+Our findings revealed that women were more likely to vote for Biden compared to Trump.
+This finding is consistent with broader trends in political behaviour as gender has previously shown to influence voting preferences (cite).
+Biden’s democratic policies were more convincing to women and will be discussed in @sec-discussion.
+Another major finding was that Black, Asian, Hispanic, and Middle Eastern individuals are also more likely to vote for Biden.
+Biden has been widely known to run on a more progressive platform, giving importance on racial justice and addressing racial disparities.
+Our finding emphasizes that people from diverse and ethnic backgrounds are more inclined to vote for progressive representatives.
+Our final major finding was that non-gun owners are also more likely to vote for Biden.
+Gun owners often have strong opinions on gun control policies.
+Our finding shows that those who do not possess guns favored Biden’s gun control and other firearm-related policies, as opposed to Trump’s policies.
 
-The survey data for this paper is from the Harvard Dataverse Repository.
-The Cooperative Election Study is a sample survey that consists of pre-election and post-election questions.
-They provide a guide and a dataset for numerous years.
-It is also open to the public.
-[add more]
-
+This paper was structured to communicate our mode, results, and analysis to the reader in a strategic manner.
 Data for this analysis and the different applications used will be further introduced in @sec-data.
@@
 # Data {#sec-data}
+
+## Sources and Methodology
+
+The survey data for this paper is from the Harvard Dataverse Repository.
+The Cooperative Election Study is a sample survey in the Harvard Dataverse Repository that consists of pre-election and post-election questions with 61000 answers.
+They provide a guide and a dataset for numerous years.
+It is also open to the public.
+For this paper, we utilized the data from the 2020 Presidential Elections(cite).
+The study was done to gain an understanding of “how Americans view Congress and hold their representatives accountable during elections, how they voted and their electoral experiences, and how their behaviour and experiences vary with political geography and social context” (cite).
+The outcome of the 2020 election and every presidential election usually changes the direction of the country and demonstrates the perspectives of the majority of the electorate.
+Thus, this dataset was able to give us some clarity on perceptions of the electorate.
+There were similar datasets available, but we chose this one due to relevance to our research question.
+The questionnaire asked various questions on race, political party preference, voting preference, ownership of a gun and have numerous multiple choice answers available for each question.
+It was also easily accessible and we were able to adapt and adjust code from Telling Stories with Data to suit our needs [@rohan].
+
+Our paper aims to address the following two questions using the data: (1) Does gender and race play a role in political preference? (2) Do those that own a gun favor one candidate over the other?
+
+The language and environment used for this analysis is R [@citeR], alongside the tidyverse [@tidyverse], dataverse [@dataverse], arrow [@arrow], rstanarm [@rstanarm], knitr [@knitr], here [@here], ggplot2 [@ggplot2], scales [@scales], modelsummary [@modelsummary], and marginaleffects [@marginaleffects] packages.
+
+## Variables and Measurement
 
@@
 
-## Sources and Collection
+In this paper, we measured key variables related to voter behaviour.
+From the dataset, we selected `gender`, `race`, `gunown`.
+In Figure 1, we can see support for candidate based on gender and race.
+For gender, we considered two options, male or female.
+For race, we considered 8 options, White, Black, Hispanic, Asian, Native American, Middle Eastern, two or more races, and other.
+Note that only one selection was allowed in the original study for both gender and race.
+Figure 2 illustrates the support for each candidate based on gun ownership status, with four options.
+Individuals who personally own a gun, who don’t personally own one, but someone in their household owns a gun, no one in the household owns a gun, and not sure.
+We also selected the variable that recorded whether an individual voted for Biden or Trump in the 2020 Presidential election, `CC20_410`.
+This allowed us to establish a binary outcome variable for the probability of voting Biden or Trump in our model in @sec-model.
+The variables measured here allowed us to identify variation in political attitudes and voting patterns.
 
-For this paper, we utilized the data from the 2020 Presidential Elections.
-We selected, gender, race, and whether an individual or a family member in the household owns a gun.
-We also selected the variable that recorded whether an individual voted for Biden or Trump.
-The broader context of the dataset is the political landscape of the US during the election cycle of 2020.
-The outcome of the 2020 election and every presidential election usually changes the direction of the country and demonstrates the perspectives of the majority of the electorate.
-There may have been similar dataset available, but we chose this one due to relevance to our research question.
-The questionnaire asked various questions on race, political party preference, voting preference, ownership of a gun and have numerous multiple choice answers available.
-Thus, this survey data provided us with the information on the demographic characteristics we wanted to consider.
-
-Our paper aims to address the following two questions: (1) Does gender and race play a role in political preference? (2) Do those that own a gun favour one candidate over the other?
-
-## Methodology
-
-## Variables
-
-## Measurement
+Data was downloaded and cleaned to suit our needs and during the data cleaning process, we made sure to only include participants who were adults registered to vote.
+We also removed rows with NA values for `gunown`.
+This was due to the fact that we wanted to see individuals who answered all the questions for our variables of interest.
+We almost limited the participants to those who explicitly stated they voted for Biden or Trump.
+Other options were available to participants such as “I did not vote” and “not sure”.
 

### 2024-03-16 paper/paper.qmd
Update paper with citations

@@
 thanks: "Code and data are available at: [https://github.com/atn-ly/political_support_in_the_us](https://github.com/atn-ly/political_support_in_the_us)"
-date: today
+date: 16 March 2024
 date-format: long
@@
 Moreover, Former President Trump had controversial and career-damaging events.
-During the 2016 elections, Trump even faced accusations regarding election rigging (cite).
-Before the 2020 elections, Trump was charged with misconduct while in office(cite).
-However, the senate acquitted him of those misconduct charges.
+Before the 2020 elections, Trump was charged with misconduct while in office but the senate acquitted him of those misconduct charges [@impeachment].
 Thus, both Biden and Trump faced have previously criticism from the electorate.
@@
 
-Based on our model and analysis, we found that certain demographic groups were mor include to support either candidate.
+Based on our model and analysis, we found that certain demographic groups were more include to support either candidate.
 Our findings revealed that women were more likely to vote for Biden compared to Trump.
-This finding is consistent with broader trends in political behaviour as gender has previously shown to influence voting preferences (cite).
+This finding is consistent with broader trends in political behavior, as gender has previously shown to influence voting preferences.
 Biden’s democratic policies were more convincing to women and will be discussed in @sec-discussion.
@@
 
-This paper was structured to communicate our mode, results, and analysis to the reader in a strategic manner.
+This paper was structured to communicate our model, results, and analysis to the reader in a strategic manner.
 Data for this analysis and the different applications used will be further introduced in @sec-data.
@@
 It is also open to the public.
-For this paper, we utilized the data from the 2020 Presidential Elections(cite).
-The study was done to gain an understanding of “how Americans view Congress and hold their representatives accountable during elections, how they voted and their electoral experiences, and how their behaviour and experiences vary with political geography and social context” (cite).
+For this paper, we utilized the data from the 2020 Presidential Elections [@CES].
+The study was done to gain an understanding of “how Americans view Congress and hold their representatives accountable during elections, how they voted and their electoral experiences, and how their behaviour and experiences vary with political geography and social context” [@CES].
 The outcome of the 2020 election and every presidential election usually changes the direction of the country and demonstrates the perspectives of the majority of the electorate.
@@
 The questionnaire asked various questions on race, political party preference, voting preference, ownership of a gun and have numerous multiple choice answers available for each question.
-It was also easily accessible and we were able to adapt and adjust code from Telling Stories with Data to suit our needs [@rohan].
+It was also easily accessible and we were able to adapt and adjust code from Telling Stories with Data [@rohan] to suit our needs.
 
@@
 For gender, we considered two options, male or female.
-For race, we considered 8 options, White, Black, Hispanic, Asian, Native American, Middle Eastern, two or more races, and other.
+For race, we considered eight options, White, Black, Hispanic, Asian, Native American, Middle Eastern, two or more races, and other.
 Note that only one selection was allowed in the original study for both gender and race.
@@
 This was due to the fact that we wanted to see individuals who answered all the questions for our variables of interest.
-We almost limited the participants to those who explicitly stated they voted for Biden or Trump.
-Other options were available to participants such as “I did not vote” and “not sure”.
+We also limited the participants to those who explicitly stated they voted for Biden or Trump.
+Initially, other options were available to participants such as “I did not vote” and “not sure”.
 

### 2024-03-16 paper/paper.qmd
Update paper with citations

@@
 
-The language and environment used for this analysis is R [@citeR], alongside the tidyverse [@tidyverse], dataverse [@dataverse], arrow [@arrow], rstanarm [@rstanarm], knitr [@knitr], here [@here], ggplot2 [@ggplot2], scales [@scales], modelsummary [@modelsummary], and marginaleffects [@marginaleffects] packages.
+The language and environment used for this analysis is `R` [@citeR], alongside the `tidyverse` [@tidyverse], `dataverse` [@dataverse], `arrow` [@arrow], `rstanarm` [@rstanarm], `knitr` [@knitr], `here` [@here], `ggplot2` [@ggplot2], `scales` [@scales], `modelsummary` [@modelsummary], and `marginaleffects` [@marginaleffects] packages.
 

### 2024-03-16 scripts/03-test_data.R
Update 03-test_data.R

@@ -5,7 +5,8 @@
 # Date: 15 March 2024
 # Contact: renfrew.aoieong@mail.utoronto.ca, rahma.binthmohammad@mail.utoronto.ca, annatn.ly@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Run the "02-data.cleaning.R" script to generate ces2020_analysis_data.parquet
+# Pre-requisites: Run the "02-data.cleaning.R" script to generate ces2020_analysis_data.parquet and "04-model.R to generate political_preferences.rds
+# Note: To check dataset without checking model, comment out lines 20-21 and 61-82
 
 
 #### Workspace setup ####

### 2024-03-17 paper/paper.qmd
Updata subsections of the data section of paper

@@
 
-## Sources and Methodology
+## Sources
 
@@
 
+## Methodology
+
 The language and environment used for this analysis is `R` [@citeR], alongside the `tidyverse` [@tidyverse], `dataverse` [@dataverse], `arrow` [@arrow], `rstanarm` [@rstanarm], `knitr` [@knitr], `here` [@here], `ggplot2` [@ggplot2], `scales` [@scales], `modelsummary` [@modelsummary], and `marginaleffects` [@marginaleffects] packages.
 
-## Variables and Measurement
+## Variables
 
@@
 
-In this paper, we measured key variables related to voter behaviour.
-From the dataset, we selected `gender`, `race`, `gunown`.
-In Figure 1, we can see support for candidate based on gender and race.
-For gender, we considered two options, male or female.
+From the dataset, we selected `gender`, `race`, `gunown` from the pre-election questionnaire.
+In @fig-race-gender, we can see support for candidate based on gender and race.
+For gender, we were given two options in the study, male or female.
 For race, we considered eight options, White, Black, Hispanic, Asian, Native American, Middle Eastern, two or more races, and other.
 Note that only one selection was allowed in the original study for both gender and race.
-Figure 2 illustrates the support for each candidate based on gun ownership status, with four options.
+@fig-gun-ownership-gender illustrates the support for each candidate based on gun ownership status, with four options.
 Individuals who personally own a gun, who don’t personally own one, but someone in their household owns a gun, no one in the household owns a gun, and not sure.
-We also selected the variable that recorded whether an individual voted for Biden or Trump in the 2020 Presidential election, `CC20_410`.
+We also selected a variable from the post-election questionnaire, the variable that recorded whether an individual voted for Biden or Trump in the 2020 Presidential election, `CC20_410`.
 This allowed us to establish a binary outcome variable for the probability of voting Biden or Trump in our model in @sec-model.
-The variables measured here allowed us to identify variation in political attitudes and voting patterns.
 
@@
 Initially, other options were available to participants such as “I did not vote” and “not sure”.
+
+## Measurement
 

### 2024-03-17 paper/paper.qmd
Add measurement section

@@
 ## Measurement
+
+We will now discuss how the variables, `gender`, `race`, and `gunown`, were measured and how they became an entry in our dataset.
+Each variable was associated with a related question, and the self-reported response of each individual from the online questionnaire was recorded in the dataset.
+Some notes on measurement error and missing data are required for each variable.
+
+For the `gender` variable, individuals may choose to identify with diverse gender identities that are beyond the traditional categorization.
+The number of voters for genders other than male or female were not collected by CES, so we have missing data that we could not include in our analysis.
+From “Counting the Countless,” this can be considered a result of what Keyes’ refers to as the “administrative violence” that reinforces the gender binary [@counting].
+
+For the `race` variable, participants were asked to select their ethnicity from one of the predetermined categories.
+Since this was an internet survey, this minimized the measurement error compared to other modes such as by a telephone interview conducted by a human.
+In the past, race has been of concern when an enumerator fills out the survey form on behalf of a respondent since it may limit the extent to which they describe their political belief [@rohan].
+
+For the `gunown` variable, participants were asked about their individual and household’s gun ownership status.
+Since the survey conductors collect Personally Identifying Information (PII), such as name, email address, and postcode [@yougov], respondents could be less likely to disclose their gun ownership status as it is a federal crime to possess illegal firearms.
 

### 2024-03-17 paper/paper.qmd
update wording

--- 2024-03-16 paper/paper.qmd
@@
 @fig-model-gun-owner shows the estimated probability that an individual supports Biden given their gun ownership status.
-We can see that Personal gun ownership and the Not Sure if anyone in the household owns a gun categories show a lower support of Biden, while the No One in the household and the Someone else in the household owns a gun categories show a higher trend of support for Biden.
+We can see that the "Personal" and "Not sure" categories show a lower support of Biden, while the "No One in the household" and the "Someone in the household" categories show a higher trend of support for Biden.
 There appears to be an expected cluster of blue dots higher up and red dots lower down, which supports the accuracy of our model.

### 2024-03-17 paper/paper.qmd
Fix grammatical errors

@@
 The presidential race between the two was of significant interest to statisticians and polling experts due to its potential impact on polling models and election forecasting.
-To provide some background, President Biden ran on his merit as a former Vice President of eight years and had faced lots of criticism during his time in office.
-Moreover, Former President Trump had controversial and career-damaging events.
-Before the 2020 elections, Trump was charged with misconduct while in office but the senate acquitted him of those misconduct charges [@impeachment].
-Thus, both Biden and Trump faced have previously criticism from the electorate.
-An interesting question to consider is despite these, what characteristics did the voters have?
+To provide some background, as former Vice President, President Biden had faced lots of scrutiny during his time in office.
+Likewise, former President Trump had controversial and career-damaging events.
+Before the 2020 elections, Trump was charged with misconduct while in office, but was acquitted by the Senate [@impeachment].
+Thus, both Biden and Trump have previously faced criticism from the electorate.
+An interesting question to consider is what characteristics did the voters have?
 Do they tell us a story on the type of people that voted for Biden and the type that voted for Trump?
@@
 For our paper, we used a logistic regression model to estimate the likelihood of individuals voting for Biden or Trump, given certain characteristics.
-We use a binary outcome variable and three predictor variables.
+We used a binary outcome variable and three predictor variables.
 Our estimand is the probability that an individual voted for Biden or Trump based on three characteristics; their gender, race, and ownership of a gun.
 The reason for this choice will be further examined in @sec-model.
-This data driven approach to elections can bring insight to voter characteristics and the importance of survey data and knowing the electorate.
+This approach to elections can bring insight to voter characteristics and the importance of survey data and knowing the electorate.
 We sought to produce results that could be used to predict voter behavior in the future.
 
-Based on our model and analysis, we found that certain demographic groups were more include to support either candidate.
-Our findings revealed that women were more likely to vote for Biden compared to Trump.
+Based on our model and analysis, our findings revealed that women were more likely to vote for Biden compared to Trump.
 This finding is consistent with broader trends in political behavior, as gender has previously shown to influence voting preferences.
 Biden’s democratic policies were more convincing to women and will be discussed in @sec-discussion.
-Another major finding was that Black, Asian, Hispanic, and Middle Eastern individuals are also more likely to vote for Biden.
+Another major finding was that Black, Asian, Hispanic, and Middle Eastern individuals were also more likely to vote for Biden.
 Biden has been widely known to run on a more progressive platform, giving importance on racial justice and addressing racial disparities.
 Our finding emphasizes that people from diverse and ethnic backgrounds are more inclined to vote for progressive representatives.
-Our final major finding was that non-gun owners are also more likely to vote for Biden.
+Our final major finding was that non-gun owners were also more likely to vote for Biden.
 Gun owners often have strong opinions on gun control policies.
@@
 
-From the dataset, we selected `gender`, `race`, `gunown` from the pre-election questionnaire.
-In @fig-race-gender, we can see support for candidate based on gender and race.
+From the dataset, we selected `gender`, `race`, and `gunown` from the pre-election questionnaire.
+In @fig-race-gender, we can see support for the candidate based on gender and race.
 For gender, we were given two options in the study, male or female.
-For race, we considered eight options, White, Black, Hispanic, Asian, Native American, Middle Eastern, two or more races, and other.
+For race, we considered eight options: White, Black, Hispanic, Asian, Native American, Middle Eastern, two or more races, and other.
 Note that only one selection was allowed in the original study for both gender and race.

### 2024-03-17 paper/paper.qmd
Fixed minor grammatical errors for final submission

@@
 Using data obtained during the fall of 2020 from the Cooperative Election Study, we investigated if a person is likely to vote for Biden, based only on knowing their gender, race, and gun ownership status.
-The results of our analysis using a Bayesian logistic regression model show that female, Black, Asian, Hispanic, Middle Eastern, and non-gun owning individuals are more likely to support Biden.
+The results of our analysis, using a Bayesian logistic regression model, show that female, Black, Asian, Hispanic, Middle Eastern, and non-gun owning individuals are more likely to support Biden.
 Our discovery highlights the importance of women’s issues, racial justice, and gun policy on American voters."
@@
 
-We will now discuss how the variables, `gender`, `race`, and `gunown`, were measured and how they became an entry in our dataset.
-Each variable was associated with a related question, and the self-reported response of each individual from the online questionnaire was recorded in the dataset.
+We will now discuss how the variables `gender`, `race`, and `gunown` were measured and how they became an entry in our dataset.
+Each variable was associated with a related question, and the self-reported responses of each individual from the online questionnaire were recorded in the dataset.
 Some notes on measurement error and missing data are required for each variable.
@@
 
-Given that Biden and Trump are far apart on women’s issues, racial justice, and gun policy, we  chose “gender”, “race”, and “gunown” to be the predictor variables in our model.
+Given that Biden and Trump are far apart on women’s issues, racial justice, and gun policy, we  chose `gender`, `race`, and `gunown` to be the predictor variables in our model.
 First, since we are interested in forecasting who a respondent is likely to vote for, we used “gender” instead of “gender_post” because it is representative of all adult Americans registered to vote rather than only adult Americans who voted.
@@
 
-# Additional data details
-
 # Model details {#sec-model-details}

### 2024-03-17 paper/paper.qmd
Fixed minor typos for final submission

@@
 Given that Biden and Trump are far apart on women’s issues, racial justice, and gun policy, we  chose `gender`, `race`, and `gunown` to be the predictor variables in our model.
-First, since we are interested in forecasting who a respondent is likely to vote for, we used “gender” instead of “gender_post” because it is representative of all adult Americans registered to vote rather than only adult Americans who voted.
-Second, we used “race” instead of variables by country or region to affirm that race is a social construct with no biological foundation.
-Third, we used “gunown” instead of variables about gun control because it captures broader opinions rather than specific policy proposals.
+First, since we are interested in forecasting who a respondent is likely to vote for, we used `gender` instead of `gender_post` because it is representative of all adult Americans registered to vote rather than only adult Americans who voted.
+Second, we used `race` instead of variables by country or region to affirm that race is a social construct with no biological foundation.
+Third, we used `gunown` instead of variables about gun control because it captures broader opinions rather than specific policy proposals.
 After exploring the data in the previous section, we found differences in political preference based on these features and that it would be of interest to investigate further.
